---
title: "Learning BNs - ALARM synthetic 1000"
output: html_notebook
---

Learning of BNs from augmented datasets created via deep generative models.

Get the libraries
```{r}
library(bnlearn)
library(dplyr)
library(Rgraphviz)
```

Import the original data
```{r}
df <- read.csv("/alarm/alarm.csv")
df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)], 
                                       as.factor)
df[sapply(df, is.logical)] <- lapply(df[sapply(df, is.logical)], 
                                       as.factor)
```

Original BN
```{r}
modelstring = paste0("[HISTORY|LVFAILURE][CVP|LVEDVOLUME][PCWP|LVEDVOLUME][HYPOVOLEMIA][LVEDVOLUME|HYPOVOLEMIA:LVFAILURE][LVFAILURE]",
  "[STROKEVOLUME|HYPOVOLEMIA:LVFAILURE][ERRLOWOUTPUT][HRBP|ERRLOWOUTPUT:HR][HREKG|ERRCAUTER:HR][ERRCAUTER][HRSAT|ERRCAUTER:HR][INSUFFANESTH]",
  "[ANAPHYLAXIS][TPR|ANAPHYLAXIS][EXPCO2|ARTCO2:VENTLUNG][KINKEDTUBE][MINVOL|INTUBATION:VENTLUNG][FIO2][PVSAT|FIO2:VENTALV]",
  "[SAO2|PVSAT:SHUNT][PAP|PULMEMBOLUS][PULMEMBOLUS][SHUNT|INTUBATION:PULMEMBOLUS][INTUBATION][PRESS|INTUBATION:KINKEDTUBE:VENTTUBE][DISCONNECT]",
  "[MINVOLSET][VENTMACH|MINVOLSET][VENTTUBE|DISCONNECT:VENTMACH][VENTLUNG|INTUBATION:KINKEDTUBE:VENTTUBE][VENTALV|INTUBATION:VENTLUNG]",
  "[ARTCO2|VENTALV][CATECHOL|ARTCO2:INSUFFANESTH:SAO2:TPR][HR|CATECHOL][CO|HR:STROKEVOLUME][BP|CO:TPR]")
original = model2network(modelstring)
graphviz.plot(original, shape = "ellipse")
fitted_original = bn.fit(original, df, method = "bayes")

distance <- 0
bayes_factor <- 0
bics <- score(original, df, type = "bic")
kullback <- 0
```

Get the sample 1,000 dataset
```{r}
stratified_1000 = read.csv("/alarm/Samples/stratified_1000.csv")
stratified_1000[sapply(stratified_1000, is.character)] <- 
  lapply(stratified_1000[sapply(stratified_1000, is.character)],as.factor)
stratified_1000[sapply(stratified_1000, is.logical)] <- 
  lapply(stratified_1000[sapply(stratified_1000, is.logical)],as.factor)

summary(stratified_1000)
```


Import the synthetic 1000 data
```{r}
synthetic_1000 = read.csv("/alarm/Synthetic/TVAE/alarm_1000_syn.csv", colClasses = "character")
synthetic_1000[synthetic_1000 == "False"] <- "FALSE"
synthetic_1000[synthetic_1000 == "True"] <- "TRUE"

synthetic_1000[sapply(synthetic_1000, is.character)] <- 
  lapply(synthetic_1000[sapply(synthetic_1000, is.character)],as.factor)

summary(synthetic_1000)
```

# Get 1000*2 synthetic samples
```{r}
set.seed(1)
synthetic_1000_2 <- synthetic_1000 %>%
  group_by(MINVOLSET) %>%
  sample_frac(size=0.1)
synthetic_1000_2 <- as.data.frame(synthetic_1000_2)
summary(synthetic_1000_2)
```

Perform Union on the datasets
```{r}
synthetic_1000_u_2 <- rbind(synthetic_1000_2, stratified_1000)
summary(synthetic_1000_u_2)
```

Learn the BN structure with the synthetic data from stratified_1000_2
```{r}
BN_synth_1000_2 <- tabu(synthetic_1000_u_2)
fitted_synth_1000_2 = bn.fit(original, synthetic_1000_u_2, method = "bayes")
graphviz.plot(BN_synth_1000_2)

distance <- c(distance, shd(original, BN_synth_1000_2, debug = TRUE))
bayes_factor <- c(bayes_factor, BF(original, BN_synth_1000_2, df, score = "bds"))
bics <- c(bics,score(BN_synth_1000_2, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_synth_1000_2))
```

# Get 1000*5 synthetic samples
```{r}
set.seed(1)
synthetic_1000_5 <- synthetic_1000 %>%
  group_by(MINVOLSET) %>%
  sample_frac(size=0.25)
synthetic_1000_5 <- as.data.frame(synthetic_1000_5)
summary(synthetic_1000_5)
```

Perform Union on the datasets
```{r}
synthetic_1000_u_5 <- rbind(synthetic_1000_5, stratified_1000)
summary(synthetic_1000_u_5)
```

Learn the BN structure with the synthetic data from stratified_1000_5
```{r}
BN_synth_1000_5 <- tabu(synthetic_1000_u_5)
fitted_synth_1000_5 = bn.fit(original, synthetic_1000_u_5, method = "bayes")
graphviz.plot(BN_synth_1000_5)

distance <- c(distance, shd(original, BN_synth_1000_5, debug = TRUE))
bayes_factor <- c(bayes_factor, BF(original, BN_synth_1000_5, df, score = "bds"))
bics <- c(bics,score(BN_synth_1000_5, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_synth_1000_5))
```

# Get 1000*10 synthetic samples
```{r}
set.seed(1)
synthetic_1000_10 <- synthetic_1000 %>%
  group_by(MINVOLSET) %>%
  sample_frac(size=0.50)
synthetic_1000_10 <- as.data.frame(synthetic_1000_10)
summary(synthetic_1000_10)
```

Perform Union on the datasets
```{r}
synthetic_1000_u_10 <- rbind(synthetic_1000_10, stratified_1000)
summary(synthetic_1000_u_10)
```

Learn the BN structure with the synthetic data from stratified_1000_10
```{r}
BN_synth_1000_10 <- tabu(synthetic_1000_u_10)
fitted_synth_1000_10 = bn.fit(original, synthetic_1000_u_10, method = "bayes")
graphviz.plot(BN_synth_1000_10)

distance <- c(distance, shd(original, BN_synth_1000_10, debug = TRUE))
bayes_factor <- c(bayes_factor, BF(original, BN_synth_1000_10, df, score = "bds"))
bics <- c(bics,score(BN_synth_1000_10, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_synth_1000_10))
```

# Get 1000*20 synthetic samples
```{r}
set.seed(1)
synthetic_1000_20 <- synthetic_1000
summary(synthetic_1000_20)
```

Perform Union on the datasets
```{r}
synthetic_1000_u_20 <- rbind(synthetic_1000_20, stratified_1000)
summary(synthetic_1000_u_20)
```

Learn the BN structure with the synthetic data from stratified_500_20
```{r}
BN_synth_1000_20 <- tabu(synthetic_1000_u_20)
fitted_synth_1000_20 = bn.fit(original, synthetic_1000_u_20, method = "bayes")
graphviz.plot(BN_synth_1000_20)

distance <- c(distance, shd(original, BN_synth_1000_20, debug = TRUE))
bayes_factor <- c(bayes_factor, BF(original, BN_synth_1000_20, df, score = "bds"))
bics <- c(bics,score(BN_synth_1000_20, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_synth_1000_20))
```

# Metrics
```{r}
matrix_empty <- matrix(nrow = 5, ncol = 1)
metrics_table = data.frame(matrix_empty)
metrics_table$bics <- bics
metrics_table$bf <- bayes_factor
metrics_table$shd <- distance
metrics_table$kullback <- kullback
metrics_table <- metrics_table[,-c(1)]
row.names(metrics_table) <- c("Full", "Sample*2", "Sample*5", "Sample*10", "Sample*20")
metrics_table
```