---
title: "Learning BNs - prm synthetic 100"
output: html_notebook
---

Learning of BNs from augmented datasets created via deep generative models.

Get the libraries
```{r}
library(bnlearn)
library(dplyr)
library(Rgraphviz)
library(Metrics)
```

Import the original data
```{r}
df = read.csv("/prm/prm.csv")
df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)], 
                                       as.factor)
```

Original BN
```{r}
dag <- empty.graph(nodes = c("Schedule","Budget","Risk_1","Risk_2","Risk_3",
                             "Project_Risk","Complexity","Type","Overrun"))
arc.set <- matrix(c("Schedule","Risk_1",
                    "Budget", "Risk_2",
                    "Risk_1", "Risk_2",
                    "Risk_2", "Risk_3",
                    "Complexity", "Risk_3",
                    "Risk_1", "Project_Risk",
                    "Risk_2", "Project_Risk",
                    "Risk_3", "Project_Risk",
                    "Project_Risk", "Overrun",
                    "Complexity", "Overrun",
                    "Type", "Overrun"),
                  byrow = TRUE, ncol = 2,
                  dimnames = list(NULL, c("from","to")))
arcs(dag) <- arc.set
original = dag

fitted_original = read.net("/prm/prm.net")

distance <- 0
bayes_factor <- 0
bics <- score(original, df, type = "bic")
kullback <- 0
```

Get the sample 100 dataset
```{r}
stratified_100 = read.csv("/prm/Samples/stratified_100.csv")
stratified_100[sapply(stratified_100, is.character)] <- 
  lapply(stratified_100[sapply(stratified_100, is.character)],as.factor)
summary(stratified_100)
```

Import the synthetic 100 data
```{r}
synthetic_100 = read.csv("/prm/Synthetic/CTGAN/prm_100_syn.csv")
synthetic_100[sapply(synthetic_100, is.character)] <- 
  lapply(synthetic_100[sapply(synthetic_100, is.character)],as.factor)
summary(synthetic_100)
```

# Get 100*2 synthetic samples
```{r}
set.seed(1)
synthetic_100_2 <- synthetic_100 %>%
  group_by(Type) %>%
  sample_frac(size=0.1)
synthetic_100_2 <- as.data.frame(synthetic_100_2)
summary(synthetic_100_2)
```

Perform Union on the datasets
```{r}
synthetic_100_u_2 <- rbind(synthetic_100_2, stratified_100)
summary(synthetic_100_u_2)
write.csv(synthetic_100_u_2, "/prm/Augdata/CTGAN/aug_100_2.csv", row.names=FALSE)
```

Learn the BN structure with the synthetic data from stratified_100_2
```{r}
BN_synth_100_2 <- tabu(synthetic_100_u_2)
fitted_synth_100_2 = bn.fit(original, synthetic_100_u_2, method = "bayes")
graphviz.plot(BN_synth_100_2)

distance <- c(distance, shd(original, BN_synth_100_2, debug = TRUE))
bayes_factor <- c(bayes_factor, BF(original, BN_synth_100_2, df, score = "bds"))
bics <- c(bics,score(BN_synth_100_2, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_synth_100_2))
```

# Get 100*5 synthetic samples
```{r}
set.seed(1)
synthetic_100_5 <- synthetic_100 %>%
  group_by(Type) %>%
  sample_frac(size=0.25)
synthetic_100_5 <- as.data.frame(synthetic_100_5)
summary(synthetic_100_5)
```

Perform Union on the datasets
```{r}
synthetic_100_u_5 <- rbind(synthetic_100_5, stratified_100)
summary(synthetic_100_u_5)
write.csv(synthetic_100_u_5, "/prm/Augdata/CTGAN/aug_100_5.csv", row.names=FALSE)
```

Learn the BN structure with the synthetic data from stratified_100_5
```{r}
BN_synth_100_5 <- tabu(synthetic_100_u_5)
fitted_synth_100_5 = bn.fit(original, synthetic_100_u_5, method = "bayes")
graphviz.plot(BN_synth_100_5)

distance <- c(distance, shd(original, BN_synth_100_5, debug = TRUE))
bayes_factor <- c(bayes_factor, BF(original, BN_synth_100_5, df, score = "bds"))
bics <- c(bics,score(BN_synth_100_5, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_synth_100_5))
```


# Get 100*10 synthetic samples
```{r}
set.seed(1)
synthetic_100_10 <- synthetic_100 %>%
  group_by(Type) %>%
  sample_frac(size=0.5)
synthetic_100_10 <- as.data.frame(synthetic_100_10)
summary(synthetic_100_10)
```

Union the datasets
```{r}
synthetic_100_u_10 <- rbind(synthetic_100_10, stratified_100)
summary(synthetic_100_u_10)
write.csv(synthetic_100_u_10, "/prm/Augdata/CTGAN/aug_100_10.csv", row.names=FALSE)
```

Learn the BN structure with the synthetic data from stratified_100_u_10
```{r}
BN_synth_100_10 <- tabu(synthetic_100_u_10)
fitted_synth_100_10 = bn.fit(original, synthetic_100_u_10, method = "bayes")
graphviz.plot(BN_synth_100_10)

distance <- c(distance,shd(original, BN_synth_100_10, debug = TRUE))
bayes_factor <- c(bayes_factor,BF(original, BN_synth_100_10, df, score = "bds"))
bics <- c(bics,score(BN_synth_100_10, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_synth_100_10))
```

# Get 100*20 synthetic samples
```{r}
set.seed(1)
synthetic_100_20 <- synthetic_100
summary(synthetic_100_20)
```


Union the datasets
```{r}
synthetic_100_u_20 <- rbind(synthetic_100_20, stratified_100)
summary(synthetic_100_u_20)
write.csv(synthetic_100_u_20, "/prm/Augdata/CTGAN/aug_100_20.csv", row.names=FALSE)
```


Learn the BN structure with the synthetic data from stratified_100_u_20
```{r}
BN_synth_100_20 <- tabu(synthetic_100_u_20)
fitted_synth_100_20 = bn.fit(original, synthetic_100_u_20, method = "bayes")
graphviz.plot(BN_synth_100_20)

distance <- c(distance,shd(original, BN_synth_100_20, debug = TRUE))
bayes_factor <- c(bayes_factor,BF(original, BN_synth_100_20, df, score = "bds"))
bics <- c(bics, score(BN_synth_100_20, df, type = "bic"))

kullback <- c(kullback, KL(fitted_original, fitted_synth_100_20))
```

# Metrics
```{r}
matrix_empty <- matrix(nrow = 5, ncol = 1)
metrics_table = data.frame(matrix_empty)
metrics_table$bics <- bics
metrics_table$bf <- bayes_factor
metrics_table$shd <- distance
metrics_table$kullback <- kullback
metrics_table <- metrics_table[,-c(1)]
row.names(metrics_table) <- c("Full", "Sample*2", "Sample*5", "Sample*10", "Sample*20")
metrics_table
```

Read the test dataset
```{r}
test_data = read.csv("/Users/camilodavid/Library/CloudStorage/OneDrive-Personal/Tohoku U/PhD/Papers/4rd Paper/Code/prm/test.csv")
test_data[sapply(test_data, is.character)] <- lapply(test_data[sapply(test_data, is.character)], 
                                       as.factor)
summary(test_data)
```

Calculate the MAP for Project Risk with synthetic_100_5
```{r}
fitted_bn_synth_100_5 <- bn.fit(BN_synth_100_5, synthetic_100_u_5, method = "bayes")
predicted_100_5 = predict(fitted_bn_synth_100_5, node = "Overrun", data = test_data)
accuracy(test_data$Overrun, predicted_100_5)
```

Calculate the MAP for Project Risk with synthetic_100_10
```{r}
fitted_bn_synth_100_10 <- bn.fit(BN_synth_100_10, synthetic_100_u_10, method = "bayes")
predicted_100_10 = predict(fitted_bn_synth_100_10, node = "Overrun", data = test_data)
accuracy(test_data$Overrun, predicted_100_10)
```

Calculate the MAP for Project Risk with synthetic_100_20
```{r}
fitted_bn_synth_100_20 <- bn.fit(BN_synth_100_20, synthetic_100_u_20, method = "bayes")
predicted_100_20 = predict(fitted_bn_synth_100_20, node = "Overrun", data = test_data)
accuracy(test_data$Overrun, predicted_100_20)
```