---
title: "Learning BNs - SPORTS synthetic 500"
output: html_notebook
---

Learning of BNs from augmented datasets created via deep generative models.

Get the libraries
```{r}
library(bnlearn)
library(dplyr)
library(Rgraphviz)
```

Import the original data
```{r}
df <- read.csv("/sports/sports.CSV")
df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)], 
                                       as.factor)
```
Original BN
```{r}
modelstring = paste0("[RDlevel][possession|RDlevel][HTshots|RDlevel:possession][ATshots|RDlevel:possession][HTshotOnTarget|RDlevel:HTshots][ATshotsOnTarget|RDlevel:ATshots][HTgoals|HTshotOnTarget:RDlevel][ATgoals|ATshotsOnTarget:RDlevel][HDA|ATgoals:HTgoals]")
original = model2network(modelstring)
graphviz.plot(original, shape = "ellipse")

fitted_original = bn.fit(original, df, method = "bayes")

distance <- 0
bayes_factor <- 0
bics <- score(original, df, type = "bic")
kullback <- 0
```

Get the sample 500 dataset
```{r}
stratified_500 = read.csv("/sports/Samples/stratified_500.csv")
stratified_500[sapply(stratified_500, is.character)] <- 
  lapply(stratified_500[sapply(stratified_500, is.character)],as.factor)
summary(stratified_500)
```

Import the synthetic 500 data
```{r}
synthetic_500 = read.csv("/sports/Synthetic/Mostly/sports_500_syn.csv")
synthetic_500[sapply(synthetic_500, is.character)] <- 
  lapply(synthetic_500[sapply(synthetic_500, is.character)],as.factor)
summary(synthetic_500)
```

# Get 500*2 synthetic samples
```{r}
set.seed(1)
synthetic_500_2 <- synthetic_500 %>%
  group_by(RDlevel) %>%
  sample_frac(size=0.1)
synthetic_500_2 <- as.data.frame(synthetic_500_2)
summary(synthetic_500_2)
```

Perform Union on the datasets
```{r}
synthetic_500_u_2 <- rbind(synthetic_500_2, stratified_500)
summary(synthetic_500_u_2)
write.csv(synthetic_500_u_2, "/sports/Augdata/Mostly/aug_500_2.csv", row.names=FALSE)
```

Learn the BN structure with the synthetic data from stratified_500_2
```{r}
BN_synth_500_2 <- tabu(synthetic_500_u_2)
fitted_synth_500_2 = bn.fit(original, synthetic_500_u_2, method = "bayes")
graphviz.plot(BN_synth_500_2)

distance <- c(distance, shd(original, BN_synth_500_2, debug = TRUE))
bayes_factor <- c(bayes_factor, BF(original, BN_synth_500_2, df, score = "bds"))
bics <- c(bics,score(BN_synth_500_2, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_synth_500_2))
```

# Get 500*5 synthetic samples
```{r}
set.seed(1)
synthetic_500_5 <- synthetic_500 %>%
  group_by(RDlevel) %>%
  sample_frac(size=0.25)
synthetic_500_5 <- as.data.frame(synthetic_500_5)
summary(synthetic_500_5)
```

Perform Union on the datasets
```{r}
synthetic_500_u_5 <- rbind(synthetic_500_5, stratified_500)
summary(synthetic_500_u_5)
write.csv(synthetic_500_u_5, "/sports/Augdata/Mostly/aug_500_5.csv", row.names=FALSE)
```

Learn the BN structure with the synthetic data from stratified_500_5
```{r}
BN_synth_500_5 <- tabu(synthetic_500_u_5)
fitted_synth_500_5 = bn.fit(original, synthetic_500_u_5, method = "bayes")
graphviz.plot(BN_synth_500_5)

distance <- c(distance, shd(original, BN_synth_500_5, debug = TRUE))
bayes_factor <- c(bayes_factor, BF(original, BN_synth_500_5, df, score = "bds"))
bics <- c(bics,score(BN_synth_500_5, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_synth_500_5))
```

# Get 500*10 synthetic samples
```{r}
set.seed(1)
synthetic_500_10 <- synthetic_500 %>%
  group_by(RDlevel) %>%
  sample_frac(size=0.50)
synthetic_500_10 <- as.data.frame(synthetic_500_10)
summary(synthetic_500_10)
```

Perform Union on the datasets
```{r}
synthetic_500_u_10 <- rbind(synthetic_500_10, stratified_500)
summary(synthetic_500_u_10)
write.csv(synthetic_500_u_10, "/sports/Augdata/Mostly/aug_500_10.csv", row.names=FALSE)
```

Learn the BN structure with the synthetic data from stratified_500_10
```{r}
BN_synth_500_10 <- tabu(synthetic_500_u_10)
fitted_synth_500_10 = bn.fit(original, synthetic_500_u_10, method = "bayes")
graphviz.plot(BN_synth_500_10)

distance <- c(distance, shd(original, BN_synth_500_10, debug = TRUE))
bayes_factor <- c(bayes_factor, BF(original, BN_synth_500_10, df, score = "bds"))
bics <- c(bics,score(BN_synth_500_10, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_synth_500_10))
```

# Get 500*20 synthetic samples
```{r}
set.seed(1)
synthetic_500_20 <- synthetic_500
summary(synthetic_500_20)
```

Perform Union on the datasets
```{r}
synthetic_500_u_20 <- rbind(synthetic_500_20, stratified_500)
summary(synthetic_500_u_20)
write.csv(synthetic_500_u_20, "/sports/Augdata/Mostly/aug_500_20.csv", row.names=FALSE)
```

Learn the BN structure with the synthetic data from stratified_500_20
```{r}
BN_synth_500_20 <- tabu(synthetic_500_u_20)
fitted_synth_500_20 = bn.fit(original, synthetic_500_u_20, method = "bayes")
graphviz.plot(BN_synth_500_20)

distance <- c(distance, shd(original, BN_synth_500_20, debug = TRUE))
bayes_factor <- c(bayes_factor, BF(original, BN_synth_500_20, df, score = "bds"))
bics <- c(bics,score(BN_synth_500_20, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_synth_500_20))
```

# Metrics
```{r}
matrix_empty <- matrix(nrow = 5, ncol = 1)
metrics_table = data.frame(matrix_empty)
metrics_table$bics <- bics
metrics_table$bf <- bayes_factor
metrics_table$shd <- distance
metrics_table$kullback <- kullback
metrics_table <- metrics_table[,-c(1)]
row.names(metrics_table) <- c("Full", "Sample*2", "Sample*5", "Sample*10", "Sample*20")
metrics_table
```

