---
title: "Learning BNs - ALARM"
output: html_notebook
---
Beinlich, I. A., Suermondt, H. J., Chavez, R. M., and Cooper, G. F. (1989). The ALARM Monitoring
System: A Case Study with Two Probabilistic Inference Techniques for Belief Networks. In
Proceedings of the 2nd European Conference on Artificial Intelligence in Medicine, pp. 247–256.

Get the libraries
```{r}
library(bnlearn)
library(dplyr)
library(Rgraphviz)
```

Import the data
```{r}
df <- read.csv("/alarm/alarm.csv")
df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)], 
                                       as.factor)
df[sapply(df, is.logical)] <- lapply(df[sapply(df, is.logical)], 
                                       as.factor)
summary(df)
```

Original BN
```{r}
modelstring = paste0("[HISTORY|LVFAILURE][CVP|LVEDVOLUME][PCWP|LVEDVOLUME][HYPOVOLEMIA][LVEDVOLUME|HYPOVOLEMIA:LVFAILURE][LVFAILURE]",
  "[STROKEVOLUME|HYPOVOLEMIA:LVFAILURE][ERRLOWOUTPUT][HRBP|ERRLOWOUTPUT:HR][HREKG|ERRCAUTER:HR][ERRCAUTER][HRSAT|ERRCAUTER:HR][INSUFFANESTH]",
  "[ANAPHYLAXIS][TPR|ANAPHYLAXIS][EXPCO2|ARTCO2:VENTLUNG][KINKEDTUBE][MINVOL|INTUBATION:VENTLUNG][FIO2][PVSAT|FIO2:VENTALV]",
  "[SAO2|PVSAT:SHUNT][PAP|PULMEMBOLUS][PULMEMBOLUS][SHUNT|INTUBATION:PULMEMBOLUS][INTUBATION][PRESS|INTUBATION:KINKEDTUBE:VENTTUBE][DISCONNECT]",
  "[MINVOLSET][VENTMACH|MINVOLSET][VENTTUBE|DISCONNECT:VENTMACH][VENTLUNG|INTUBATION:KINKEDTUBE:VENTTUBE][VENTALV|INTUBATION:VENTLUNG]",
  "[ARTCO2|VENTALV][CATECHOL|ARTCO2:INSUFFANESTH:SAO2:TPR][HR|CATECHOL][CO|HR:STROKEVOLUME][BP|CO:TPR]")
original = model2network(modelstring)
graphviz.plot(original, shape = "ellipse")
score(original, df, type = "bic")
fitted_original = bn.fit(original, df, method = "bayes")
```
sample 1000 data points
```{r}
set.seed(1)
stratified_1000 <- df %>%
  group_by(MINVOLSET) %>%
  sample_frac(size=0.01)
stratified_1000 <- as.data.frame(stratified_1000)
summary(stratified_1000)
write.csv(stratified_1000, "/alarm/Samples/stratified_1000.csv", row.names=FALSE)
```

sample 500 data points
```{r}
set.seed(1)
stratified_500 <- stratified_1000 %>%
  group_by(MINVOLSET) %>%
  sample_frac(size=0.5)
stratified_500 <- as.data.frame(stratified_500)
summary(stratified_500)
write.csv(stratified_500, "/alarm/Samples/stratified_500.csv", row.names=FALSE)
```

sample 100 data points
```{r}
set.seed(1)
stratified_100 <- stratified_500 %>%
  group_by(MINVOLSET) %>%
  sample_frac(size=0.2)
stratified_100 <- as.data.frame(stratified_100)
summary(stratified_100)
write.csv(stratified_100, "/alarm/Samples/stratified_100.csv", row.names=FALSE)
```

Learn the BN structure with full data
```{r}
BN_full <- tabu(df)
fitted_full = bn.fit(original, df, method = "bayes")
graphviz.plot(BN_full)

distance <- shd(original, BN_full, debug = TRUE)
bayes_factor <- BF(original, BN_full, df, score = "bds")
bics <- score(BN_full, df, type = "bic")

kullback <- KL(fitted_original, fitted_full)
```

Learn the BN structure with 100 data points
```{r}
BN_100 <- tabu(stratified_100)
fitted_100 <- bn.fit(original, stratified_100, method = "bayes")
graphviz.plot(BN_100)

distance <- c(distance,shd(original, BN_100, debug = TRUE))
bayes_factor <- c(bayes_factor,BF(original, BN_100, df, score = "bds"))
bics <- c(bics,score(BN_100, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_100))
```

Learn the BN structure with 500 data points
```{r}
BN_500 <- tabu(stratified_500)
fitted_500 <- bn.fit(original, stratified_500, method = "bayes")
graphviz.plot(BN_500)

distance <- c(distance,shd(original, BN_500, debug = TRUE))
bayes_factor <- c(bayes_factor,BF(original, BN_500, df, score = "bds"))
bics <- c(bics,score(BN_500, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_500))
```

Learn the BN structure with 1,000 data points
```{r}
BN_1000 <- tabu(stratified_1000)
fitted_1000 <- bn.fit(original, stratified_1000, method = "bayes")
graphviz.plot(BN_1000)

distance <- c(distance,shd(original, BN_1000, debug = TRUE))
bayes_factor <- c(bayes_factor,BF(original, BN_1000, df, score = "bds"))
bics <- c(bics,score(BN_1000, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_1000))
```


# Metrics
```{r}
matrix_empty <- matrix(nrow = 4, ncol = 1)
metrics_table = data.frame(matrix_empty) 
metrics_table$bics <- bics
metrics_table$bf <- bayes_factor
metrics_table$shd <- distance
metrics_table$kullback <- kullback
metrics_table <- metrics_table[,-c(1)]
row.names(metrics_table) <- c("Full", "Sample100", "Sample500", "Sample1000")
metrics_table
```