---
title: "Learning BNs - INSURANCE"
output: html_notebook
---
Insurance is a network for evaluating car insurance risks.
Binder J, Koller D, Russell S, Kanazawa K (1997). "Adaptive Probabilistic Networks with Hidden Variables". Machine Learning, 29(2–3):213–244.

Get the libraries
```{r}
library(bnlearn)
library(dplyr)
library(Rgraphviz)
```

Import the data
```{r}
data("insurance")
df <- insurance
summary(df)
```

Original BN
```{r}
modelstring = paste0("[Age][Mileage][SocioEcon|Age][GoodStudent|Age:SocioEcon]",
  "[RiskAversion|Age:SocioEcon][OtherCar|SocioEcon][VehicleYear|SocioEcon:RiskAversion]",
  "[MakeModel|SocioEcon:RiskAversion][SeniorTrain|Age:RiskAversion]",
  "[HomeBase|SocioEcon:RiskAversion][AntiTheft|SocioEcon:RiskAversion]",
  "[RuggedAuto|VehicleYear:MakeModel][Antilock|VehicleYear:MakeModel]",
  "[DrivingSkill|Age:SeniorTrain][CarValue|VehicleYear:MakeModel:Mileage]",
  "[Airbag|VehicleYear:MakeModel][DrivQuality|RiskAversion:DrivingSkill]",
  "[Theft|CarValue:HomeBase:AntiTheft][Cushioning|RuggedAuto:Airbag]",
  "[DrivHist|RiskAversion:DrivingSkill][Accident|DrivQuality:Mileage:Antilock]",
  "[ThisCarDam|RuggedAuto:Accident][OtherCarCost|RuggedAuto:Accident]",
  "[MedCost|Age:Accident:Cushioning][ILiCost|Accident]",
  "[ThisCarCost|ThisCarDam:Theft:CarValue][PropCost|ThisCarCost:OtherCarCost]")
original = model2network(modelstring)
graphviz.plot(original, shape = "ellipse")
score(original, df, type = "bic")
fitted_original = bn.fit(original, df, method = "bayes")
```
sample 1000 data points
```{r}
set.seed(1)
stratified_1000 <- df %>%
  group_by(Age) %>%
  sample_frac(size=0.05)
stratified_1000 <- as.data.frame(stratified_1000)
summary(stratified_1000)
write.csv(stratified_1000, "/insurance/Samples/stratified_1000.csv", row.names=FALSE)
```

sample 500 data points
```{r}
set.seed(1)
stratified_500 <- stratified_1000 %>%
  group_by(Age) %>%
  sample_frac(size=0.5)
stratified_500 <- as.data.frame(stratified_500)
summary(stratified_500)
write.csv(stratified_500, "/insurance/Samples/stratified_500.csv", row.names=FALSE)
```

sample 100 data points
```{r}
set.seed(1)
stratified_100 <- stratified_500 %>%
  group_by(Age) %>%
  sample_frac(size=0.2)
stratified_100 <- as.data.frame(stratified_100)
summary(stratified_100)
write.csv(stratified_100, "/insurance/Samples/stratified_100.csv", row.names=FALSE)
```

Learn the BN structure with full data
```{r}
BN_full <- tabu(df)
fitted_full = bn.fit(original, df, method = "bayes")
graphviz.plot(BN_full)

distance <- shd(original, BN_full, debug = TRUE)
bayes_factor <- BF(original, BN_full, df, score = "bds")
bics <- score(BN_full, df, type = "bic")

kullback <- KL(fitted_original, fitted_full)
```

Learn the BN structure with 100 data points
```{r}
BN_100 <- tabu(stratified_100)
fitted_100 <- bn.fit(original, stratified_100, method = "bayes")
graphviz.plot(BN_100)

distance <- c(distance,shd(original, BN_100, debug = TRUE))
bayes_factor <- c(bayes_factor,BF(original, BN_100, df, score = "bds"))
bics <- c(bics,score(BN_100, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_100))
```

Learn the BN structure with 500 data points
```{r}
BN_500 <- tabu(stratified_500)
fitted_500 <- bn.fit(original, stratified_500, method = "bayes")
graphviz.plot(BN_500)

distance <- c(distance,shd(original, BN_500, debug = TRUE))
bayes_factor <- c(bayes_factor,BF(original, BN_500, df, score = "bds"))
bics <- c(bics,score(BN_500, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_500))
```

Learn the BN structure with 1,000 data points
```{r}
BN_1000 <- tabu(stratified_1000)
fitted_1000 <- bn.fit(original, stratified_1000, method = "bayes")
graphviz.plot(BN_1000)

distance <- c(distance,shd(original, BN_1000, debug = TRUE))
bayes_factor <- c(bayes_factor,BF(original, BN_1000, df, score = "bds"))
bics <- c(bics,score(BN_1000, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_1000))
```


# Metrics
```{r}
matrix_empty <- matrix(nrow = 4, ncol = 1)
metrics_table = data.frame(matrix_empty) 
metrics_table$bics <- bics
metrics_table$bf <- bayes_factor
metrics_table$shd <- distance
metrics_table$kullback <- kullback
metrics_table <- metrics_table[,-c(1)]
row.names(metrics_table) <- c("Full", "Sample100", "Sample500", "Sample1000")
metrics_table
```