---
title: "prm Project Risk Management"
output: html_notebook
---

Load the libraries
```{r}
library(bnlearn)
library(Rgraphviz)
library(dplyr)
```

Creating the Structure

```{r}
dag <- empty.graph(nodes = c("Schedule","Budget","Risk_1","Risk_2","Risk_3",
                             "Project_Risk","Complexity","Type","Overrun"))
arc.set <- matrix(c("Schedule","Risk_1",
                    "Budget", "Risk_2",
                    "Risk_1", "Risk_2",
                    "Risk_2", "Risk_3",
                    "Complexity", "Risk_3",
                    "Risk_1", "Project_Risk",
                    "Risk_2", "Project_Risk",
                    "Risk_3", "Project_Risk",
                    "Project_Risk", "Overrun",
                    "Complexity", "Overrun",
                    "Type", "Overrun"),
                  byrow = TRUE, ncol = 2,
                  dimnames = list(NULL, c("from","to")))
arcs(dag) <- arc.set
graphviz.plot(dag, shape = "ellipse")
```

Assign the level to each node
```{r}
Schedule.lv <- c("short","large")
Budget.lv <- c("small","large")
Risk_1.lv <- c("yes","no")
Risk_2.lv <- c("yes","no")
Risk_3.lv <- c("yes","no")
Project_Risk.lv <- c("low","medium", "high")
Complexity.lv <- c("easy","hard")
Type.lv <- c("A","B")
Overrun.lv <- c("acceptable","unacceptable")
```

Assign the probabilities to each node
```{r}
Schedule.prob <- array(c(0.78,0.22), dim = 2, dimnames = list(Schedule = Schedule.lv))
Budget.prob <- array(c(0.67,0.33), dim = 2, dimnames = list(Budget = Budget.lv))
Complexity.prob <- array(c(0.8,0.2), dim = 2, dimnames = list(Complexity = Complexity.lv))
Type.prob <- array(c(0.9,0.1), dim = 2, dimnames = list(Type = Type.lv))

Risk_1.prob <- array(c(0.6,0.4,0.5,0.5), dim = c(2,2), 
                     dimnames = list(Risk_1 = Risk_1.lv,
                                     Schedule = Schedule.lv))

Risk_2.prob <- array(c(0.6,0.4,0.8,0.2,0.2,0.8,0.3,0.7), dim = c(2,2,2), 
                     dimnames = list(Risk_2 = Risk_2.lv,
                                     Budget = Budget.lv,
                                     Risk_1 = Risk_1.lv
                                     ))

Risk_3.prob <- array(c(0.5,0.5,0.9,0.1,0.1,0.9,0.4,0.6), dim = c(2,2,2), 
                     dimnames = list(Risk_3 = Risk_3.lv,
                                     Complexity = Complexity.lv,
                                     Risk_2 = Risk_2.lv
                                     ))

Project_Risk.prob <- array(c(0.05,0.1,0.85,
                             0.3,0.4,0.3,
                             0.1,0.2,0.7,
                             0.5,0.3,0.2,
                             0.1,0.3,0.6,
                             0.5,0.4,0.1,
                             0.3,0.4,0.3,
                             0.85,0.1,0.05), dim = c(3,2,2,2), 
                     dimnames = list(Project_Risk = Project_Risk.lv,
                                     Risk_3 = Risk_3.lv,
                                     Risk_2 = Risk_2.lv,
                                     Risk_1 = Risk_1.lv
                                     ))

Overrun.prob <- array(c(0.95,0.05,
                        0.90,0.10,
                        0.80,0.20,
                        0.75,0.25,
                        0.60,0.40,
                        0.55,0.45,
                        0.45,0.55,
                        0.40,0.60,
                        0.25,0.75,
                        0.20,0.80,
                        0.10,0.90,
                        0.05,0.95), dim = c(2,2,2,3), 
                     dimnames = list(Overrun = Overrun.lv,
                                     Type = Type.lv,
                                     Complexity = Complexity.lv,
                                     Project_Risk = Project_Risk.lv
                                     ))
```

Create the fitted BN
```{r}
cpt <- list(Schedule=Schedule.prob,
            Budget=Budget.prob,
            Complexity=Complexity.prob,
            Type=Type.prob,
            Risk_1=Risk_1.prob,
            Risk_2=Risk_2.prob,
            Risk_3=Risk_3.prob,
            Project_Risk=Project_Risk.prob,
            Overrun=Overrun.prob)

bn <- custom.fit(dag,cpt)
write.net("/prm/prm.net",bn)
```

Create the data
```{r}
set.seed(1)
df <- rbn(bn, 100000)
write.csv(df, "/prm/prm.csv", row.names=FALSE)
df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)], 
                                       as.factor)
test_data <- rbn(bn, 1000)
write.csv(test_data, "/prm/test.csv", row.names=FALSE)
summary(df)
```

Original BN
```{r}
original = dag
score(original, df, type = "bic")
fitted_original = bn
```

sample 1000 data points
```{r}
set.seed(1)
stratified_1000 <- df %>%
  group_by(Type) %>%
  sample_frac(size=0.01)
stratified_1000 <- as.data.frame(stratified_1000)
summary(stratified_1000)
write.csv(stratified_1000, "/prm/Samples/stratified_1000.csv", row.names=FALSE)
```

sample 500 data points
```{r}
set.seed(1)
stratified_500 <- stratified_1000 %>%
  group_by(Type) %>%
  sample_frac(size=0.5)
stratified_500 <- as.data.frame(stratified_500)
summary(stratified_500)
write.csv(stratified_500, "/prm/Samples/stratified_500.csv", row.names=FALSE)
```

sample 100 data points
```{r}
set.seed(1)
stratified_100 <- stratified_500 %>%
  group_by(Type) %>%
  sample_frac(size=0.2)
stratified_100 <- as.data.frame(stratified_100)
summary(stratified_100)
write.csv(stratified_100, "/prm/Samples/stratified_100.csv", row.names=FALSE)
```

Learn the BN structure with full data
```{r}
BN_full <- tabu(df)
fitted_full = bn.fit(original, df, method = "bayes")
graphviz.plot(BN_full, shape = "ellipse")

distance <- shd(original, BN_full, debug = TRUE)
bayes_factor <- BF(original, BN_full, df, score = "bds")
bics <- score(BN_full, df, type = "bic")

kullback <- KL(fitted_original, fitted_full)
```

Learn the BN structure with 100 data points
```{r}
BN_100 <- tabu(stratified_100)
fitted_100 <- bn.fit(original, stratified_100, method = "bayes")
graphviz.plot(BN_100, shape = "ellipse")

distance <- c(distance,shd(original, BN_100, debug = TRUE))
bayes_factor <- c(bayes_factor,BF(original, BN_100, df, score = "bds"))
bics <- c(bics,score(BN_100, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_100))
```

Learn the BN structure with 500 data points
```{r}
BN_500 <- tabu(stratified_500)
fitted_500 <- bn.fit(original, stratified_500, method = "bayes")
graphviz.plot(BN_500, shape = "ellipse")

distance <- c(distance,shd(original, BN_500, debug = TRUE))
bayes_factor <- c(bayes_factor,BF(original, BN_500, df, score = "bds"))
bics <- c(bics,score(BN_500, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_500))
```


Learn the BN structure with 1,000 data points
```{r}
BN_1000 <- tabu(stratified_1000)
fitted_1000 <- bn.fit(original, stratified_1000, method = "bayes")
graphviz.plot(BN_1000, shape = "ellipse")

distance <- c(distance,shd(original, BN_1000, debug = TRUE))
bayes_factor <- c(bayes_factor,BF(original, BN_1000, df, score = "bds"))
bics <- c(bics,score(BN_1000, df, type = "bic"))

kullback <- c(kullback,KL(fitted_original, fitted_1000))
```


# Metrics
```{r}
matrix_empty <- matrix(nrow = 4, ncol = 1)
metrics_table = data.frame(matrix_empty) 
metrics_table$bics <- bics
metrics_table$bf <- bayes_factor
metrics_table$shd <- distance
metrics_table$kullback <- kullback
metrics_table <- metrics_table[,-c(1)]
row.names(metrics_table) <- c("Full", "Sample100", "Sample500", "Sample1000")
metrics_table
```


Calculate the MAP for Project Risk with stratified_100
```{r}
fitted_bn_100 <- bn.fit(BN_100, stratified_100, method = "bayes")
predicted_100 = predict(fitted_bn_100, node = "Overrun", data = test_data)
accuracy(test_data$Overrun, predicted_100)
```

Calculate the MAP for Project Risk with stratified_500
```{r}
fitted_bn_500 <- bn.fit(BN_500, stratified_500, method = "bayes")
predicted_500 = predict(fitted_bn_500, node = "Overrun", data = test_data)
accuracy(test_data$Overrun, predicted_500)
```

Calculate the MAP for Project Risk with stratified_1000
```{r}
fitted_bn_1000 <- bn.fit(BN_1000, stratified_1000, method = "bayes")
predicted_1000 = predict(fitted_bn_1000, node = "Overrun", data = test_data)
accuracy(test_data$Overrun, predicted_1000)
```